Pretty darn good control: when are approximate solutions better than approximate models

### 2023-06-26 manuscript/manuscript.Rmd
mini
added a bit discussion, fixed some typos

More specifically, for each parametric noise strength $\sigma_{\text{param.}}$ choice in $[0.04, 0.08, \dots, 0.2]$, we executed the following procedure:
Sample 100 choices of perturbed parameter values, where the perturbation is as follows---each parameter $P$ is perturbed to $(1+g_P)P$ where $g_P$ is a Gaussian random variable with variance $\sigma_{\text{param.}}^2$.
For each of these parameter choices, tune the constant escapement strategy and train the DRL model on the corresponding model.
Measure the average reward difference between strategies for each of these parameter choice samples using 100 replicate evaluation episodes.
Take the mean of this average difference over the 100 perturbed parameter samples.
The stability results are displayed in Fig. \textbf{TBD}.

@@

We propose an alternative approach to the problem of fishery control: to use a more expressive -- albeit more complicated -- dynamical model to guide decision making.
Furthermore, rather than computing the optimal control policy for the model (something that is impossible in practice for complex systems) we use deep reinforcement learning to obtain a ‘‘pretty darn good’’ policy.
This policy is estimated in a *model-free* setting, i.e., the agent treats the dynamical model (e.g. eq. \eqref{eq:3d model stoch}) as a black box of input-output pairs.
By not relying on precise knowledge of the model's parameter values, but rather just relying on input-output statistics, model-free approaches have gained traction in a variety of control theory settings (see, e.g., [@sato2019; @ramirez2022; @zeng2019]).

We compare deep reinforcement learning-based policies against optimal constant mortality policies. 
While the latter are MSY-inspired, they are optimized in a model-free way as well: the optimal constant mortality rate is computed empirically from simulated data.

Because of the simplicity of the constant mortality ansatz, the optimal such policy may be easily estimated (see Figure \ref{fig:tuning}).
In contrast, DRL optimizes over the more expressive -- and more complicated -- family of policies parametrized by a neural network.
Because of this, our DRL approach uses a variant of *gradient ascent* (the proximal policy optimization algorithm) to obtain a good candidate policy.

>>>>

More specifically, for each parametric noise strength $\sigma_{\text{param.}}$ choice in $[0.04, 0.08, \dots, 0.2]$, we executed the following procedure:
Sample 100 choices of perturbed parameter values, where the perturbation is as follows---each parameter $P$ is perturbed to $(1+g_P)P$ where $g_P$ is a Gaussian random variable with variance $\sigma_{\text{param.}}^2$.
For each of these parameter choices, tune the constant escapement strategy and train the DRL model on the corresponding model.
Measure the average reward difference between strategies for each of these parameter choice samples using 100 replicate evaluation episodes.
Take the mean of this average difference over the 100 perturbed parameter samples.

@@

We propose an alternative approach to the problem of fishery control: to use a more expressive -- albeit more complicated -- dynamical model to guide decision making.
Furthermore, rather than computing the optimal control policy for the model (something that is impossible in practice for complex systems) we use deep reinforcement learning to obtain a ‘‘pretty darn good’’ policy.
This policy is estimated in a *model-free* setting, i.e., the agent treats the dynamical model (e.g. eq. \eqref{eq:3d model}) as a black box of input-output pairs.
By not relying on precise knowledge of the model's parameter values, but rather just relying on input-output statistics, model-free approaches have gained traction in a variety of control theory settings (see, e.g., [@sato2019; @ramirez2022; @zeng2019]).

We compare deep reinforcement learning-based policies against classical management strategies (optimal constant mortality and optimal constant escapement policies policies). 
While the latter are inspired by the shape of optimal solutions in the single-species setting, they are optimized in a model-free way as well: e.g. the optimal constant mortality rate is computed empirically from simulated data.

Because of the simplicity of the classical policy functions, the optimal such policy may be easily estimated through a grid-search.
This is the case since these policy functions are only specified by one or two parameters (respectively in the single fishery, or two fishery cases).
In contrast, DRL optimizes over the more expressive -- and more complicated -- family of policies parametrized by a neural network.
Because of this, our DRL approach uses a variant of *gradient ascent* (the proximal policy optimization algorithm) to obtain a good candidate policy.

We showed that for sufficiently complex management scenarios -- with multiple harvested species and, possibly, time-varying parameters -- DRL-based management strategies perform significantly better than constant escapement.
This with respect to both average rewards received and conservation goals.
In this sense, an approximate solution to a more complicated and expressive model, can outperform the optimal solution of the single-species problem---even when the parameters of the single-species solution are empirically optimized.

What is more, this DRL advantage happened within a regime of parameter space where species interaction terms were much smaller than the single-species logistic growth terms.
Indeed, interaction parameters in \eqref{eq:3d model} are about an order of magnitude smaller than $r_X$ and $r_Y$.
In this situation logistic growth is a reasonable approximation of the system, and thus a constant escapement solution is a plausible control strategy.

We found that the optimal constant mortality policy surprisingly performs much better than constant escapement (Fig. \ref{fig:main}).
However, it can be observed in Fig. \ref{fig:fracmsy} that the constant mortality strategy faces a trade-off: high sustainability is achieved for sub-optimal constant mortality rates, but only at a significant decrease in the mean episode reward.
We expect that with increasing ecosystem complexity this phenomenon might become more pronounced.
We can understand this as a consequence of the rigidity of classical strategies: the simplicity of their expressions, depending only on a few parameters, means that policy optimization is constrained to a rather reduced subset of the space of possible policies.

Finally, we performed a stability analysis to ensure that the advantage of DRL-based techniques over constant escapement is a ubiquitous phenomenon and not a result of a lucky selection of parameter values.
We found (in Fig. \ref{fig:stability}) that an advantage can be observed even for relatively high-noise perturbations of the parameters---noise with a variance of 20% of the parameter values.
The aforementioned Figure summarizes the statistics of 100 parameter perturbations so that we may expect perturbations of up to 60% (i.e. three sigmas) in the parameter values to appear in these statistics.

# Old discussion
